Community-Conditioned Language Models

### 2021-06-11 paper/paper.tex
small fixes

The conditioned models also perform differently across different communities---%
even among the best models, some communities have $\IG_{M_j} < 1$,
meaning that the CCLM performs worse than the unconditioned baseline for 
messages from that community.
For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
(\cref{fig:comm-info-gain})\jp{missing label}.\footnote{
  Some of the communities with consistently high $IG_{M_j}$ 
  across all models are primarily non-English, 
  but surprisingly, not the three most extreme
  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
  the later two of which are places where people coordinate to play video games
  together. The coordination has highly structured and regular formats, which 
  are presumably conventional to these communities.}
%
The variation in information gain across communities is not 
related to perplexity, however (\cref{fig:ppl-info-gain}).

@@

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinctive its language is.
\jp{Conclusion: perplexity (linguistic complexity) is not correlated
  to how recognisable a community is. Presumably (in general) some very specific
  lexical items are characteristic of a community.}
%
\begin{table}
  \small
  \centering
  \begin{tabular}{llrrrr}
  \toprule
      & $l_c$ & $\Ppl_{M_j}$ & $\IG_{M_j}$ \\
  \midrule
  \multirow{4}{*}{LSTM}         & 0 &         0.21 &         -0.41 \\
                                & 1 &         0.19 &         -0.40 \\
                                & 2 &         0.19 &         -0.43 \\
                                & 3 &         0.06 &         -0.45 \\
  \midrule
  \multirow{4}{*}{Transformer}  & 0 &         0.19 &         -0.41 \\
                                & 1 &         0.23 &         -0.43 \\
                                & 2 &         0.16 &         -0.39 \\
                                & 3 &         0.12 &         -0.40 \\
  \bottomrule
  \end{tabular}
  \caption{ Pearson's $r$ correlation coefficient between community
    indiscernibility ($\Ind_j$) and two different predictors: CCLM
    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$). $p < 0.001$
    for all correlations except $\Ppl_{M_j}$ for the LSTM with 
  $l_c=3$ ($p=0.218$) and Transfomer with $l_c=3$ ($p=0.010$).}
  \label{tab:ind-corrs}
\end{table}

@@

Finally, we note that, even though community conditioning affects
perplexity very little overall, it provides large benefits in term of
the underlying Bayesian classifier. Indiscernibility is on average
\num{0.57} with a standard deviation of \num{0.22}. Roughly speaking, this means
that, while the linguistic variations do not allow our model to
pinpoint the originating community of a message, it can often
eliminate TODO (a percentatge?) of communities as possible origins \bn{43\%? JP: yes}
for it. Furthermore, the information gain provided by
community-conditioning is a good predictor of how specific the
community is.\jp{By empirical observation?}

@@

\begin{figure*}[t]
    \centering
  \begin{tikzpicture}
    \begin{groupplot}[%
      group style={
        group size=2 by 1,
        horizontal sep=0pt,
      },
      xmin=-1, xmax=1,
      ymin=-1, ymax=1,
      xtick={-0.5,0.5}, ytick={-0.5,0.5},
      axis x line=middle,
      axis y line=middle,
      axis line style={(-)},
      axis on top=true,
      width=0.55\textwidth,
      height=0.55\textwidth,
      scatter/classes={
        0={mark=x, blue},
        1={mark=+, orange},
        2={mark=o, green},
        3={mark=triangle, gray},
        4={mark=star, magenta},
        5={mark=diamond, red}
      },
      ]
      \nextgroupplot[
        %xlabel=$U_0$, ylabel=$U_1$,
        legend cell align=left,
        legend style={font=\tiny},
        legend entries = {
          offmychest confession relationships relationship\_advice SuicideWatch, 
          TrueReddit China europe nyc socialism,
          frugalmalefashion running MechanicAdvice personalfinance Flipping,
          darksouls skyrimmods dragonage Xcom MonsterHunter,
          amiugly Rateme ladybonersgw SchoolIdolFestival GoneWildPlus,
          facepalm CrappyDesign mildlyinfuriating cringepics oddlysatisfying,
        },
        %legend pos=outer south east
        legend to name=pcalegend
      ]
      \addplot[%
        scatter, only marks,
        mark options={thick, scale=0.75},
       scatter src=explicit symbolic,
        ] table [x={web-PCA0}, y={web-PCA1}, meta={web-cluster}] {floats/comm.csv};
      \nextgroupplot[
      %xlabel=$V_0$, ylabel=$V_0$
      ]
      \addplot[%
        scatter, only marks,
        mark options={thick, scale=0.75},
       scatter src=explicit symbolic,
        ] table [x={lstm-3-1-PCA0}, y={lstm-3-1-PCA1}, meta={web-cluster}] {floats/comm.csv};
    \end{groupplot}
  \end{tikzpicture}
  \captionof{figure}{First two components of the aligned social (left) and linguistic (right) embeddings,
    where the lingusitic embedding is taken from the LTSM with $l_c=1$.
    Correlation between these directions is given by $\sigma_0 = 53.4$ and $\sigma_1 = 35.6$.
    Colors are assigned by k-means clustering of the social embedding.
    % The legend shows the closest 5 communites to each cluster centroid.
    % JP: commented out this sentence which does not seem to refer to something visible.
    % JP: If the legend shows in the supplementary material, then the whole figure should be repeated there.

  }
  \label{fig:pca-aligned}
\end{figure*}

>>>>

The conditioned models also perform differently across different communities---%
even among the best models, some communities have $\IG_{M_j} < 1$,
meaning that the CCLM performs worse than the unconditioned baseline for 
messages from that community.
For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
(\cref{fig:comm-stratified-box}).\footnote{
  Some of the communities with consistently high $IG_{M_j}$ 
  across all models are primarily non-English, 
  but surprisingly, not the three most extreme
  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
  the later two of which are places where people coordinate to play video games
  together. The coordination has highly structured and regular formats, which 
  are presumably conventional to these communities.}
%
The variation in information gain across communities is not 
related to perplexity, however (\cref{fig:ppl-info-gain}).

@@

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinctive its language is.
%\jp{Conclusion: perplexity (linguistic complexity) is not correlated
  %to how recognisable a community is. Presumably (in general) some very specific
  %lexical items are characteristic of a community.}
%
\begin{table}
  \small
  \centering
  \begin{tabular}{llrrrr}
  \toprule
      & $l_c$ & $\Ppl_{M_j}$ & $\IG_{M_j}$ \\
  \midrule
  \multirow{4}{*}{LSTM}         & 0 &         0.21 &         -0.41 \\
                                & 1 &         0.19 &         -0.40 \\
                                & 2 &         0.19 &         -0.43 \\
                                & 3 &         0.06 &         -0.45 \\
  \midrule
  \multirow{4}{*}{Transformer}  & 0 &         0.19 &         -0.41 \\
                                & 1 &         0.23 &         -0.43 \\
                                & 2 &         0.16 &         -0.39 \\
                                & 3 &         0.12 &         -0.40 \\
  \bottomrule
  \end{tabular}
  \caption{ Pearson's $r$ correlation coefficient between community
    indiscernibility ($\Ind_j$) and two different predictors: CCLM
    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$). $p < 0.001$
    for all correlations except $\Ppl_{M_j}$ for the LSTM with 
  $l_c=3$ ($p=0.218$) and Transfomer with $l_c=3$ ($p=0.010$).}
  \label{tab:ind-corrs}
\end{table}

@@

Finally, we note that, even though community conditioning affects
perplexity very little overall, it provides large benefits in term of
the underlying Bayesian classifier. Indiscernibility is on average
\num{0.57} with a standard deviation of \num{0.22}. This means
that, while the linguistic variations do not allow our model to
pinpoint the originating community of a message, it can on average
eliminate 43\% of communities as possible origins 
for it. Furthermore, the information gain provided by
community-conditioning is a good predictor of how discernable the
community is.

@@

\begin{figure*}[t]
    \centering
  \begin{tikzpicture}
    \begin{groupplot}[%
      group style={
        group size=2 by 1,
        horizontal sep=0pt,
      },
      xmin=-1, xmax=1,
      ymin=-1, ymax=1,
      xtick={-0.5,0.5}, ytick={-0.5,0.5},
      axis x line=middle,
      axis y line=middle,
      axis line style={(-)},
      axis on top=true,
      width=0.55\textwidth,
      height=0.55\textwidth,
      scatter/classes={
        0={mark=x, blue},
        1={mark=+, orange},
        2={mark=o, green},
        3={mark=triangle, gray},
        4={mark=star, magenta},
        5={mark=diamond, red}
      },
      ]
      \nextgroupplot[
        %xlabel=$U_0$, ylabel=$U_1$,
        legend cell align=left,
        legend style={font=\tiny},
        legend entries = {
          offmychest confession relationships relationship\_advice SuicideWatch, 
          TrueReddit China europe nyc socialism,
          frugalmalefashion running MechanicAdvice personalfinance Flipping,
          darksouls skyrimmods dragonage Xcom MonsterHunter,
          amiugly Rateme ladybonersgw SchoolIdolFestival GoneWildPlus,
          facepalm CrappyDesign mildlyinfuriating cringepics oddlysatisfying,
        },
        %legend pos=outer south east
        legend to name=pcalegend
      ]
      \addplot[%
        scatter, only marks,
        mark options={thick, scale=0.75},
       scatter src=explicit symbolic,
        ] table [x={web-PCA0}, y={web-PCA1}, meta={web-cluster}] {floats/comm.csv};
      \nextgroupplot[
      %xlabel=$V_0$, ylabel=$V_0$
      ]
      \addplot[%
        scatter, only marks,
        mark options={thick, scale=0.75},
       scatter src=explicit symbolic,
        ] table [x={lstm-3-1-PCA0}, y={lstm-3-1-PCA1}, meta={web-cluster}] {floats/comm.csv};
    \end{groupplot}
  \end{tikzpicture}
  \caption{First two components of the aligned social (left) and linguistic (right) embeddings,
    where the lingusitic embedding is taken from the LTSM with $l_c=1$.
    Correlation between these directions is given by $\sigma_0 = 53.4$ and $\sigma_1 = 35.6$.
    Colors are assigned by k-means clustering of the social embedding.
    % The legend shows the closest 5 communites to each cluster centroid.
    % JP: commented out this sentence which does not seem to refer to something visible.
    % JP: If the legend shows in the supplementary material, then the whole figure should be repeated there.
  }
  \label{fig:pca-aligned}
\end{figure*}
